---
title: "Final Project"
author: "Turkmnistan"
date: '2019 11 27 '
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(devtools)
library(ggplot2)
worldcup <- read.csv("worldcup2018.csv")
row.names(worldcup) <- worldcup$Countries
worldcup2 <- worldcup[, -1]
#worldcup2
worldcup2$Possession <- worldcup2$Possession/100
summary(worldcup2)

worldcup2 <- worldcup2[,-7]

#num_df <- worldcup2[, -10]
num_df <- worldcup2[, -9]
apply(num_df, 2, var) #Shots has the highest variance

#PCA
worldcup_pca <- prcomp(num_df, scale = TRUE)
summary(worldcup_pca)
print(worldcup_pca)

#Scree
screeplot(worldcup_pca, main = "", col = "green", type = "lines", pch = 1, npcs = length(worldcup_pca$sdev))

worldcup_pca$rotation <- -worldcup_pca$rotation
worldcup_pca$x <- -worldcup_pca$x
biplot(worldcup_pca)
```

##Biplot pair
```{r}
pairs(worldcup_pca$x[,1:4],
panel=function(x,y,...){
  text(x,y,cex=1,abbreviate(rownames(worldcup2)),col=c('red','green','blue'))
  #bvbox(cbind(x,y),add=TRUE)
})
```

## same as biplot
```{r fig.height= 6, fig,width= 35}
library(ggfortify)
autoplot(worldcup_pca, data = worldcup2, colour = 'Result',
         label = TRUE, label.size = 5, 
         loadings = TRUE, loadings.colour = 'black',
         loadings.label = TRUE, 
         loadings.label.size = 10, 
         loadings.label.colour = "black") + 
  theme(legend.text = element_text(size = 16), 
        legend.title = element_text(size = 24), 
        axis.title = element_text(size = 14))
```

### put this
```{r}
library(psych)
pairs.panels(worldcup2[,-9], cex.cor = 2)

win <- worldcup2[worldcup2$Result == "W", ]
draw <- worldcup2[worldcup2$Result == "D", ]
lose <-worldcup2[worldcup2$Result == "L", ]

pairs.panels(win[,-9], cex.cor = 2)
pairs.panels(draw[,-9], cex.cor = 2)
pairs.panels(lose[,-9], cex.cor = 2)

#cor(win)
#cor(draw)
#cor(lose)

```

```{r}
## Seems useless
worldcup2$Result <- as.numeric(as.factor(worldcup2$Result))
plot(worldcup2$Result, worldcup_pca$x[,1])
cor(worldcup2$Result, worldcup_pca$x[,1])

yv <- predict(worldcup_pca)[, 1]
yv2 <- predict(worldcup_pca)[, 2]
par(mfrow = c(1,2))
plot(worldcup2$Result, yv, pch = 16, xlab = "biomass", ylab = "PC 1", col = "red")
plot(worldcup2$Result, yv2, pch = 16, xlab = "soil pH", ylab = "PC 2", col = "blue")
```

##Mosaic
```{r}
mosaicplot(~ Shots.OT + Goals, data = worldcup2, color = c("green","red","blue")) #Lose, Draw, Win
```

## star plot
```{r}
stars(num_df, key.loc = c(15, 2), cex = 0.8, draw.segments = T)

stars(win, key.loc = c(15, 2), cex = 0.8, draw.segments = T)
stars(lose, key.loc = c(15, 2), cex = 0.8, draw.segments = T)
stars(draw, key.loc = c(15, 2), cex = 0.8, draw.segments = T)
```

## Heatmap + Dendogram
```{r}
bball <- as.matrix(num_df)

heatmap(bball) ## light color higher value. needs scaling
heatmap(bball, scale = "column", Colv = NA) #scaled
```

## Dendogram with cutting lines
```{r}
#wc_dist = dist(worldcup2[,-9])
#plot(cs <- hclust(wc_dist,method = "single"),main = "Single Linkage")

#plot(cc <- hclust(wc_dist,method = "complete"), main = "Complete Linkage")
#plot(ca <- hclust(wc_dist,method = "average"), main = "Average Linkage")


HC <- hclust(wc_dist, "average")
myDend <-  as.dendrogram(HC) 

par(mar=c(7.5,4,2,2))
plot(myDend, main = "Average Linkage")

k = 3
n = nrow(worldcup2)
MidPoint = (HC$height[n-k] + HC$height[n-k+1]) / 2
abline(h = MidPoint, lty=2)

```

 Now we run into the issue of choosing the "correct" amount of clusters, that is, where should we cut the dendogram? One method of determining this could be to look at how many groups a model-based clustering technique chooses, such as maximum-likelihood based clustering. However, upon further analysis, the model-based clustering chooses only 1 cluster, which does not help.
```{r,echo=F}
##Mclust
library(mclust)
mclus <- Mclust(worldcup2[,-9])
mclus$classification
```
 
## PCA Cluster. a plot of each team listed by their cluster
```{r,echo=F}

cc <- hclust(wc_dist,method = "complete")
##cc <- hclust(wc_dist,method = "single")
##cc <- hclust(wc_dist,method = "average")


clus_pca = princomp(worldcup2[,-9],cor=TRUE)
xlim = range(clus_pca$scores[,1])

par(mfrow = c(1,1))
group1 <- worldcup2$Result == 'W' 
group2 <- worldcup2$Result == 'D' 
group3 <- worldcup2$Result == 'L' 
lab = cutree(cc,k = 3);
wc3 = worldcup2

plot(clus_pca$scores[,1:2],type='n',xlim=xlim,ylim=xlim, main = "PCA Cluster")
text(clus_pca$scores[,1][group1],clus_pca$scores[,2][group1],labels=lab[group1],cex=1.5,col='blue')
text(clus_pca$scores[,1][group2],clus_pca$scores[,2][group2],labels=lab[group2],cex=1.5,col='red')
text(clus_pca$scores[,1][group3],clus_pca$scores[,2][group3],labels=lab[group3],cex=1.5,col='green')
legend("topleft",legend=c("Winners","Drawers","Losers"),cex=.75,
       col=c("blue","red","green"),pch='|')
```